---
output:
  html_document: 
    keep_md: yes
  pdf_document: default
---
# Reproducible Research Week 2: Peer Assessment 1

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}


# Default: no printings at all unless it will be overrided by a chunk

# Set working directory

setwd("Z:/USER/Axel/Fortbildung/Data Analyst/R/50 Course/050 Reproducible Research/week 2/90 Programs/Assessment 1")

# load needed libraries

library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data
##### 1. Load the data 
###### Unzip the data if not present and load via read.csv()
```{r unzip and load, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activityData <- read.csv('activity.csv')
```
###### The variables included in this dataset are:
```{r Variables, results='markup', warning=TRUE, message=FALSE, echo=FALSE}
names (activityData)
rows <-nrow (activityData)
```

- steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
- date: The date on which the measurement was taken in YYYY-MM-DD format
- interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17.568 observations in this dataset.



##### 2. Transformation not needed or be included in the further chunks

-----

## What is mean total number of steps taken per day?
```{r Calculate Mean}
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
```

##### 1. Make a histogram of the total number of steps taken each day
```{r Histogram total numers by day}
qplot(stepsByDay, xlab='Steps per day', ylab='Numbers', binwidth=1000)
```

##### 2. Calculate and report the mean and median total number of steps taken per day
```{r Calculate mean and median}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)
```
* Mean: `r stepsByDayMean`
* Median:  `r stepsByDayMedian`

-----

## What is the average daily activity pattern?
```{r Calculate Mean and Median}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```

##### 1. Make a time series plot
```{r Time series plot}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("interval") +
    ylab("average number") 
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r Maximum nuber of steps by Intervall}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```

* Most Steps at: `r timeMostSteps`

----

## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r}
numMissingValues <- length(which(is.na(activityData$steps)))
```

* Number of missing values: `r numMissingValues`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r Impute Missing}
activityDataImputed <- activityData
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
```


##### 4. Make a histogram of the total number of steps taken each day 
```{r Histogram of Steps by day}
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Steps per day', ylab='Numbers', binwidth=1000)
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r Report Histogramm}
stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
```
* Mean (Imputed): `r stepsByDayMeanImputed`
* Median (Imputed):  `r stepsByDayMedianImputed`


----

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r Subsetting Data}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### 2. Make a panel plot containing a time series plot

```{r Plot data}
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("interval") + 
    ylab("numbers")
```





